### 2015-05-14 background.tex
hacks in bg

Example~\ref{fig:data_race_example} shows a racy entry point that in the generic\_nvram Linux char driver;
our \whoop (see \S\ref{whoop}) was able to find this race automatically 
(see \S\ref{evaluation}). The driver can invoke the \texttt{nvram\_llseek} entry point concurrently from two threads $T_1$ and $T_2$, while passing the same \texttt{file} object as a parameter. However, this can lead to multiple possible data races, because $T_1$ and $T_2$ can access the \texttt{f\_pos} field of the \texttt{file} object in arbitrary order.

\begin{figure}[t]
\begin{lstlisting}
static loff_t nvram_llseek(struct file *file,
    loff_t offset, int origin) {
  switch (origin) {
    case 0: break;
    case 1: offset += file->f_pos; break; // racy
    case 2: offset += nvram_len; break;
    default: offset = -1;
  }
  if (offset < 0)
    return -EINVAL;
  file->f_pos = offset; // racy
  return file->f_pos; // racy
}
\end{lstlisting}
\caption{Example of data races in the generic\_nvram Linux char driver}
\label{fig:data_race_example}
\end{figure}

@@

Lock-based synchronization is a powerful tool for preventing races in drivers, but careless use of locks has many pitfalls~\cite{sutter2005software}. Coarse-grained locking can hurt performance, as it limits the opportunity for concurrency. On the other hand, fine-grained locking can easily lead to deadlocks. Although the Linux kernel provides sophisticated lock-free synchronization mechanisms~\cite{corbet2005linux}, such as atomic read-modify-write operations, in this work we focus on locks as they are widely used.\footnote{We treat lock-free operations by not regarding them as providing any protection between threads. This is sound, but can lead to false alarm race reports.}

>>>>

Figure~\ref{fig:data_race_example} shows a racy entry point, \texttt{nvram\_llseek}, in the generic\_nvram Linux driver. The driver can invoke the entry point concurrently from two threads $T_1$ and $T_2$, with the same \texttt{file} object as a parameter. However, this can lead to multiple possible data races, because $T_1$ and $T_2$ can access the \texttt{f\_pos} field of \texttt{file} in arbitrary order. Our tool, \whoop (see \S\ref{whoop}), was able to find these races automatically (see \S\ref{evaluation}).

\begin{figure}[t]
\begin{lstlisting}
static loff_t nvram_llseek(struct file *file,
    loff_t offset, int origin) {
  switch (origin) {
    case 0: break;
    case 1: offset += file->f_pos; break; // racy
    case 2: offset += nvram_len; break;
    default: offset = -1;
  }
  if (offset < 0)
    return -EINVAL;
  file->f_pos = offset; // racy
  return file->f_pos; // racy
}
\end{lstlisting}
\caption{Racy entry point in the generic\_nvram Linux driver}
\label{fig:data_race_example}
\end{figure}

@@

Lock-based synchronization is a powerful tool for preventing races in drivers, but careless use of locks has many pitfalls~\cite{sutter2005software}. Coarse-grained locking can hurt performance, as it limits the opportunity for concurrency. On the other hand, fine-grained locking can easily lead to deadlocks. Although the Linux kernel provides sophisticated lock-free synchronization mechanisms~\cite{corbet2005linux}, such as atomic read-modify-write operations, in this work we focus on locks as they are widely used.\footnote{We treat lock-free operations soundly by not regarding them as providing any protection between threads, but this can lead to false alarm race reports.}
